### 2024-03-05 tex/trans.tex
[ correction ] first pass over chapter 4

The types of DILL are the same as the types of \name, except for the
restriction of $\oc\gr{r}$ to just $\oc\gr\omega$.
I will write the latter simply as $\oc$ when it appears in DILL\@.
I add sums and with-products to the calculus of \citet{Barber1996}, with the
obvious rules (stated fully in \cref{fig:dill2}).
These additive type formers present no additional difficulty to the translation.

\begin{figure}
  \begin{mathpar}
    \ebrule{\infer0[Int-Ax]{\gamma, A; {\cdot} \vdash A}} \and
    \ebrule{\infer0[Lin-Ax]{\gamma; A \vdash A}} \and
    \ebrule{\infer0[$I$-I]{\gamma; {\cdot} \vdash I}} \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash I}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$I$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A}
      \hypo{\gamma, \delta_2 \vdash B}
      \infer2[$\otimes$-I]{\gamma; \delta_1, \delta_2 \vdash A \otimes B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \otimes B}
      \hypo{\gamma; \delta_2, A, B \vdash C}
      \infer2[$\otimes$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A \vdash B}
      \infer1[$\multimap$-I]{\gamma; \delta \vdash A \multimap B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \multimap B}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$\multimap$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A}
      \infer1[$\oc$-I]{\gamma; \cdot \vdash \oc A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash \oc A}
      \hypo{\gamma, A; \delta_2 \vdash B}
      \infer2[$\oc$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A}
      \hypo{\gamma, \delta \vdash B}
      \infer2[$\with$-I]{\gamma; \delta \vdash A \with B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \with A_1}
      \infer1[$\with$-E$_i$]{\gamma; \delta \vdash A_i}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash 0}
      \infer1[$0$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i}
      \infer1[$\oplus$-I$_i$]{\gamma; \delta \vdash A_0 \oplus A_1}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \oplus B}
      \hypo{\gamma; \delta_2, A \vdash C}
      \hypo{\gamma; \delta_2, B \vdash C}
      \infer3[$\oplus$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
  \end{mathpar}
  \label{fig:dill2}
  \caption{The rules of DILL, extended with additive connectives}
\end{figure}

\begin{figure}
  \begin{mathpar}
    \begin{eqns}
      \mathrm{DILL} &\hookrightarrow& \name_\instDILL \\
      Y &\mapsto& \iota_Y \\
      I &\mapsto& I \\
      A \otimes B &\mapsto& A \otimes B \\
      A \multimap B &\mapsto& A \multimap B \\
      \oc A &\mapsto& \oc{\gr\omega}{A} \\
      0 &\mapsto& 0 \\
      A \oplus B &\mapsto& A \oplus B \\
      \top &\mapsto& \top \\
      A \with B &\mapsto& A \with B
    \end{eqns}
    \and
    \begin{eqns}
      \mathrm{PD} &\hookrightarrow& \name_\instPD \\
      Y &\mapsto& \iota_Y \\
      \top &\mapsto& I \\
      A \wedge B &\mapsto& A \with B \\
      A \supset B &\mapsto& A \multimap B \\
      \Box A &\mapsto& \oc{\valid}{A} \\
      \bot &\mapsto& 0 \\
      A \vee B &\mapsto& A \oplus B
    \end{eqns}
  \end{mathpar}
  \label{fig:dill-pd}
  \caption{Embedding of DILL and PD types into \name}
\end{figure}

@@

\begin{definition}\label{def:DILL-bottom-up}
  A derivation is said to be \emph{$\instDILL$-bottom-up} if only the following
  facts about addition and multiplication are used, and all proofs of
  inequalities not at leaves are by reflexivity (i.e, not using the facts that
  $\gr\omega \leq \gr0$ and $\gr\omega \leq \gr1$).

@@

The translation to and from the modal system of \citet{judgmental}
(henceforth \emph{PD}) is similar to the translation to and from DILL\@.
I present my variant of PD, again adding some common connectives, in
\cref{fig:pd}
The main difference is the algebra at which \name{} is instantiated.

\begin{figure}
  \begin{mathpar}
    \ebrule{%
      \infer0[hyp]{\gamma; \delta, A\;\mathit{true} \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[hyp*]{\gamma, A\;\mathit{valid}; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A\;\mathit{true} \vdash B\;\mathit{true}}
      \infer1[$\supset$I]{\gamma; \delta \vdash A \supset B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \supset B\;\mathit{true}}
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \infer2[$\supset$E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A\;\mathit{true}}
      \infer1[$\Box$I]{\gamma; \delta \vdash \Box A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \Box A\;\mathit{true}}
      \hypo{\gamma, A\;\mathit{valid}; \delta \vdash B\;\mathit{true}}
      \infer2[$\Box$-E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \hypo{\gamma, \delta \vdash B\;\mathit{true}}
      \infer2[$\wedge$-I]{\gamma; \delta \vdash A \wedge B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \wedge A_1\;\mathit{true}}
      \infer1[$\wedge$-E$_i$]{\gamma; \delta \vdash A_i\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \bot\;\mathit{true}}
      \infer1[$\bot$-E]{\gamma; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i\;\mathit{true}}
      \infer1[$\vee$-I$_i$]{\gamma; \delta \vdash A_0 \vee A_1\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \vee B\;\mathit{true}}
      \hypo{\gamma; \delta, A \vdash C\;\mathit{true}}
      \hypo{\gamma; \delta, B \vdash C\;\mathit{true}}
      \infer3[$\vee$-E]{\gamma; \delta \vdash C\;\mathit{true}}
    }
  \end{mathpar}
  \label{fig:pd}
  \caption{The rules of PD, extended with several standard connectives}
\end{figure}

\begin{definition}
  Let $\instPD$ denote the following semiring on the partially ordered set
  $\{\valid \triangleleft \true \triangleleft \unused\}$.
  \begin{itemize}
    \item $0 := \unused$.
    \item $+$ is the meet ($\wedge$) according to the subusaging order.
    \item $1 := \true$.
    \item
      \begin{tabular}{c|ccc}
        $*$ & $\unused$ & $\true$ & $\valid$ \\ \hline
        $\unused$ & $\unused$ & $\unused$ & $\unused$ \\
        $\true$ & $\unused$ & $\true$ & $\valid$ \\
        $\valid$ & $\unused$ & $\valid$ & $\valid$ \\
      \end{tabular}
  \end{itemize}
\end{definition}

@@

\begin{definition}\label{def:PD-bottom-up}
  A derivation is said to be \emph{$\instPD$-bottom-up} if only the following
  facts about addition and multiplication are used, and all proofs of
  inequalities not at leaves are by reflexivity.

>>>>

The types of DILL are the same as the types of \name, except for the
restriction of $\oc\gr{r}$ to just $\oc\gr\omega$.
I will write the latter simply as $\oc$ when it appears in DILL\@.
I add sums and with-products to the calculus of \citet{Barber1996}, with the
obvious rules (stated fully in \cref{fig:dill2}).
These additive type formers present no additional difficulty to the translation.
As a cosmetic difference, I assume in $\name$ a base type $\iota_Y$ for each
atomic formula $Y$ in DILL.

\begin{figure}
  \begin{mathpar}
    \ebrule{\infer0[Int-Ax]{\gamma, A; {\cdot} \vdash A}} \and
    \ebrule{\infer0[Lin-Ax]{\gamma; A \vdash A}} \and
    \ebrule{\infer0[$I$-I]{\gamma; {\cdot} \vdash I}} \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash I}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$I$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A}
      \hypo{\gamma, \delta_2 \vdash B}
      \infer2[$\otimes$-I]{\gamma; \delta_1, \delta_2 \vdash A \otimes B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \otimes B}
      \hypo{\gamma; \delta_2, A, B \vdash C}
      \infer2[$\otimes$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A \vdash B}
      \infer1[$\multimap$-I]{\gamma; \delta \vdash A \multimap B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \multimap B}
      \hypo{\gamma; \delta_2 \vdash A}
      \infer2[$\multimap$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A}
      \infer1[$\oc$-I]{\gamma; \cdot \vdash \oc A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash \oc A}
      \hypo{\gamma, A; \delta_2 \vdash B}
      \infer2[$\oc$-E]{\gamma; \delta_1, \delta_2 \vdash B}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A}
      \hypo{\gamma, \delta \vdash B}
      \infer2[$\with$-I]{\gamma; \delta \vdash A \with B}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \with A_1}
      \infer1[$\with$-E$_i$]{\gamma; \delta \vdash A_i}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash 0}
      \infer1[$0$-E]{\gamma; \delta_1, \delta_2 \vdash A}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i}
      \infer1[$\oplus$-I$_i$]{\gamma; \delta \vdash A_0 \oplus A_1}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta_1 \vdash A \oplus B}
      \hypo{\gamma; \delta_2, A \vdash C}
      \hypo{\gamma; \delta_2, B \vdash C}
      \infer3[$\oplus$-E]{\gamma; \delta_1, \delta_2 \vdash C}
    }
  \end{mathpar}
  \caption{The rules of DILL, extended with additive connectives}
  \label{fig:dill2}
\end{figure}

\begin{figure}
  \begin{mathpar}
    \begin{eqns}
      \mathrm{DILL} &\hookrightarrow& \name_\instDILL \\
      Y &\mapsto& \iota_Y \\
      I &\mapsto& I \\
      A \otimes B &\mapsto& A \otimes B \\
      A \multimap B &\mapsto& A \multimap B \\
      \oc A &\mapsto& \oc{\gr\omega}{A} \\
      0 &\mapsto& 0 \\
      A \oplus B &\mapsto& A \oplus B \\
      \top &\mapsto& \top \\
      A \with B &\mapsto& A \with B
    \end{eqns}
    \and
    \begin{eqns}
      \mathrm{PD} &\hookrightarrow& \name_\instPD \\
      Y &\mapsto& \iota_Y \\
      \top &\mapsto& I \\
      A \wedge B &\mapsto& A \with B \\
      A \supset B &\mapsto& A \multimap B \\
      \Box A &\mapsto& \oc{\valid}{A} \\
      \bot &\mapsto& 0 \\
      A \vee B &\mapsto& A \oplus B
    \end{eqns}
  \end{mathpar}
  \caption{Embedding of DILL and PD types into \name}
  \label{fig:dill-pd}
\end{figure}

@@

\begin{definition}\label{def:DILL-bottom-up}
  A derivation is said to be \emph{$\instDILL$-bottom-up} if only the facts
  about addition and multiplication listed in the tables below are used,
  and all proofs of
  inequalities not at leaves are by reflexivity (i.e, not using the facts that
  $\gr\omega \leq \gr0$ and $\gr\omega \leq \gr1$).
  In the tables, each row corresponds to a value for the left argument, each
  column corresponds to a value for the right argument, and the dash (-)
  indicates a fact that is not to be used.

@@

The translation to and from the modal system of \citet{judgmental}
(henceforth \emph{PD}) is similar to the translation to and from DILL\@.
I present my variant of PD, again adding some common connectives, in
\cref{fig:pd}.
The main difference is the algebra at which \name{} is instantiated.

\begin{figure}
  \begin{mathpar}
    \ebrule{%
      \infer0[hyp]{\gamma; \delta, A\;\mathit{true} \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[hyp*]{\gamma, A\;\mathit{valid}; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta, A\;\mathit{true} \vdash B\;\mathit{true}}
      \infer1[$\supset$I]{\gamma; \delta \vdash A \supset B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \supset B\;\mathit{true}}
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \infer2[$\supset$E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \cdot \vdash A\;\mathit{true}}
      \infer1[$\Box$I]{\gamma; \delta \vdash \Box A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \Box A\;\mathit{true}}
      \hypo{\gamma, A\;\mathit{valid}; \delta \vdash B\;\mathit{true}}
      \infer2[$\Box$-E]{\gamma; \delta \vdash B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \infer0[$\top$-I]{\gamma; \delta \vdash \top\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A\;\mathit{true}}
      \hypo{\gamma, \delta \vdash B\;\mathit{true}}
      \infer2[$\wedge$-I]{\gamma; \delta \vdash A \wedge B\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_0 \wedge A_1\;\mathit{true}}
      \infer1[$\wedge$-E$_i$]{\gamma; \delta \vdash A_i\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash \bot\;\mathit{true}}
      \infer1[$\bot$-E]{\gamma; \delta \vdash A\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A_i\;\mathit{true}}
      \infer1[$\vee$-I$_i$]{\gamma; \delta \vdash A_0 \vee A_1\;\mathit{true}}
    }
    \and
    \ebrule{%
      \hypo{\gamma; \delta \vdash A \vee B\;\mathit{true}}
      \hypo{\gamma; \delta, A \vdash C\;\mathit{true}}
      \hypo{\gamma; \delta, B \vdash C\;\mathit{true}}
      \infer3[$\vee$-E]{\gamma; \delta \vdash C\;\mathit{true}}
    }
  \end{mathpar}
  \caption{The rules of PD, extended with several standard connectives}
  \label{fig:pd}
\end{figure}

\begin{definition}
  Let $\instPD$ denote the following semiring on the partially ordered set
  $\{\valid < \true < \unused\}$.
  \begin{itemize}
    \item $0 := \unused$.
    \item $+$ is the meet ($\wedge$) according to the subusaging order.
    \item $1 := \true$.
    \item
      \begin{tabular}{c|ccc}
        $*$ & $\unused$ & $\true$ & $\valid$ \\ \hline
        $\unused$ & $\unused$ & $\unused$ & $\unused$ \\
        $\true$ & $\unused$ & $\true$ & $\valid$ \\
        $\valid$ & $\unused$ & $\valid$ & $\valid$ \\
      \end{tabular}
  \end{itemize}
\end{definition}

@@

\begin{definition}\label{def:PD-bottom-up}
  A derivation is said to be \emph{$\instPD$-bottom-up} if only the facts
  about addition and multiplication listed in the tables below are used,
  and all proofs of
  inequalities not at leaves are by reflexivity.
  In the tables, each row corresponds to a value for the left argument, each
  column corresponds to a value for the right argument, and the dash (-)
  indicates a fact that is not to be used.
